S1: Extended Derivations

### 2010-06-24 ltx/PLoS/SupportingInformation.tex
further edits to support material

\section*{Spatial Cut-off Frequency of Field Basis Functions}\label{ap:FrequencyAnalysis}
This Appendix provides the derivation of the field basis function cut-off frequency as a function of the basis function width (equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} in the main text). Applying the \textit{n}-dimensional Fourier transform \cite{Arsac1966} to a \textit{n}-dimensional Gaussian centred at the origin (defined in equation~\ref{eq:BasisFunctionAtOrigin}) yields
\begin{eqnarray}\label{eq:AppendixGaussianFT}
 \lefteqn{\boldsymbol\Phi(\boldsymbol \nu)=\int_{\mathbb{R}^n}{\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r}} \nonumber \\
&=&\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \nonumber \\
&\times& \mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right).
\end{eqnarray}
Now evaluating the integral on the RHS of equation~\ref{eq:AppendixGaussianFT} gives 
\begin{eqnarray}\label{eq:IntegralOfGaussian}
\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r&& \nonumber \\
=\left(\pi\sigma_{\phi}^2\right)^{\frac{n}{2}}.&&
\end{eqnarray}
Substituting equation~\ref{eq:IntegralOfGaussian} into equation~\ref{eq:AppendixGaussianFT} gives the scaled Gaussian
\begin{equation}\label{eq:FTofGaussian}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation} 
where 
\begin{equation}
	\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}

For $3$~dB attenuation ($1/2$ to peak power) at the cut-off frequency, $\boldsymbol\nu_c$, we need to set
\begin{eqnarray}
 \left|\boldsymbol\Phi(\boldsymbol\nu_{c})\right|^2=\frac{1}{2}\left|\boldsymbol\Phi(\mathbf 0)\right|^2,
\end{eqnarray}
since the maximum of Gaussian described in equation~\ref{eq:FTofGaussian} is at the origin. Now solving for $\sigma_{\nu}^2$ we get
\begin{equation}
 \sigma_{\nu}^2=\frac{2\boldsymbol\nu_{c}^\top \boldsymbol\nu_{c}}{\ln 2 }.
\end{equation}

>>>>

\section*{Spatial Cut-off Frequency of Field Basis Functions}\label{ap:FrequencyAnalysis}
This section gives the derivation for the field basis function spatial cut-off frequency as a function of the basis function width (equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} in the main text). For we define an \textit{n}-dimensional Gaussian  centred at the origin as 
\begin{equation}\tag{S1.7}
 \phi(\mathbf r)=\mathrm{exp}\left(-\frac{1}{\sigma^2}\mathbf r^\top\mathbf r\right).
\end{equation}
Now applying the \textit{n}-dimensional Fourier transform~\cite{Arsac1966} yields
\begin{align}\label{eq:AppendixGaussianFT}
 \boldsymbol\Phi(\boldsymbol \nu)&=\int_{\mathbb{R}^n} {\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r} \nonumber \\
&=\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \times\mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right).\tag{S1.8}
\end{align}
Now evaluating the integral on the RHS of equation~S1.8 gives 
\begin{equation}\label{eq:IntegralOfGaussian}\tag{S1.9}
\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r
=\left(\pi\sigma_{\phi}^2\right)^{\frac{n}{2}}.
\end{equation}
Substituting equation~S1.9 into equation~S1.8 gives the scaled Gaussian
\begin{equation}\label{eq:FTofGaussian}\tag{S1.10}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation} 
where 
\begin{equation}\tag{S1.11}
	\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
For $3$~dB attenuation ($1/2$ to peak power) at the cut-off frequency, $\boldsymbol\nu_c$, we need to set
\begin{equation}\tag{S1.12}
 \left|\boldsymbol\Phi(\boldsymbol\nu_{c})\right|^2=\frac{1}{2}\left|\boldsymbol\Phi(\mathbf 0)\right|^2,
\end{equation}
since the maximum of Gaussian described in equation~S1.10 is at the origin. Now solving for $\sigma_{\nu}^2$ we get
\begin{equation}\tag{S1.13}
 \sigma_{\nu}^2=\frac{2\boldsymbol\nu_{c}^\top \boldsymbol\nu_{c}}{\ln 2 }.
\end{equation}
